\documentclass[thesis-solanki.tex]{subfiles}


\begin{document}

% ---------------------------------------------------------------------------
\chapter{Quasiquotation}\label{chap:quasiquotation}
% ---------------------------------------------------------------------------

\section{Theory}
\begin{enumerate}

\item Papers
\begin{enumerate}
\item
\end{enumerate}

\item Books
\begin{enumerate}
\item
\end{enumerate}

\item Articles / Blogs / Discussions
\begin{enumerate}
\item
\end{enumerate}

\item Websites
\begin{enumerate}
\item Quasiquotation Wikipedia, \cite{wikiquasi}
\item Quasiquotation in Haskell, \cite{haskellquasi}
\end{enumerate}

\end{enumerate}


\section{Implementations}
\begin{enumerate}
\item
\end{enumerate}

\section{Miscellaneous / Possibly Related Content}
\begin{enumerate}
\item
\end{enumerate}

\section{What is Quasiquotation ?}
\begin{enumerate}
\item http://plato.stanford.edu/entries/quotation/

When language is used to attribute properties to language or otherwise theorize about it, a linguistic device is needed that ‘turns 
language on itself’. Quotation is one such device. It is our primary meta-linguistic tool.

\item http://www.thefreedictionary.com/quasi-quotation

a metalinguistic device for referring to the form of an expression containing variables without referring to the symbols for those 
variables. Thus while "not p" refers to the expression consisting of the word not followed by the letter p, the quasi-quotation \newline
$\ulcorner$ not p $\urcorner$ refers to the form of any expression consisting of the word not followed by any value of the variable p.

\item  Quasiquotation Wikipedia, \cite{wikiquasi}

Quasi-quotation or Quine quotation is a linguistic device in formal languages that facilitates rigorous and terse formulation of general rules about linguistic expressions while properly observing the use–mention distinction.

https://en.wikipedia.org/wiki/Use%E2%80%93mention_distinction
The use–mention distinction is a foundational concept of analytic philosophy,[1] according to which it is necessary to make a distinction 
between using a word (or phrase) and mentioning it


\end{enumerate}

\section{Quasiquotaion in \progLang{Haskell}}
\cite{haskellquasi, mainland2007s} 

Quasiquoting allows programmers to use custom, domain-specific syntax to construct fragments of their program. Along with 
\progLang{Haskell}'s existing support for domain specific languages, you are now free to use new syntactic forms for your EDSLs. 

Working with complex data types can impose a significant syntactic burden; extensive applications of nested data constructors are often 
required to build values of a given data type, or, worse yet, to pattern match against values.

Allow \progLang{Haskell} expressions and patterns to be constructed using domain specific, programmer-defined concrete syntax.




\end{document}
